Random Generation of Rich \\ Abstract Data Type Values

### 2019-01-31 paper/introduction.tex
Pass over intro

% Scenario of random testing
Random testing is a promising approach to finding bugs
\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
%
\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
Haskell community.
%
It requires developers to specify (i) \emph{testing properties} describing
programs expected behavior and (ii) \emph{random data generators} based on the
\emph{types} of the expected inputs (e.g., an integer, an string,
etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
%
\quickcheck then generates random test cases and reports those violating testing
properties.

% Situation with Quick Check, ADTs, tools
\quickcheck comes equipped with random generators for built-in types, while it
requires to manually write generators for user-defined ADTs.
%
Recently, there have been a proliferation of tools to automatically derive
\quickcheck generators for ADTs
\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
%
The main difference about these tools lies on the guarantees provided to ensure
\emph{the termination of the generation process} and the \emph{distribution of
  random values}.
%
Despite their differences, these tools guarantee that generated values are
\emph{well-typed}.
%
In other words, generated values follow the structure described by the
definition of the ADT.

%% How to use random generated ADTs
Well-typed ADT values are specially useful when testing programs which expect
highly structured inputs, e.g., compilers \cite{Palka11,MidtgaardJKNN17}.
%
Generating ADT values also proves fruitful when looking for vulnerabilities in
combination with fuzzers \cite{GriecoCB16,grieco2017}.
%
%% Problem
%
Despite these success stories, ADT type-definitions do not often capture all the
invariants expected from the data that they are intended to model.
%
As a result, even if random values are well-typed, they might not be built with
enough structure to penetrate into deep layers of software.

%% Our proposal
In this work, we identify two different sources of structural information that
can be statically exploited to improve the generation process of ADT values
(Section \ref{sec:sources}).
%
% In this work, we propose a novel improvement in the generation process of ADT
% values by exploiting some static information found in the codebase.
%
Then, we provide a compositional approach to capture this information into our
automatically derived random generators.
%
More specifically, we propose a generation process that is capable to consider
how programs branch based on ADTs values' patterns as well as how they get
manipulated by abstract interfaces (Section \ref{sec:hrepcont}).
%
Furthermore, we show how to predict the distribution of the \emph{expected}
numbers of ADT constructors, values fitting certain branching patterns, and
calls to interfaces.
%
For that, we extend some recent results on applying \emph{branching
  processes}\cite{gw1875}---a simple stochastic model conceived to study
population growth (Section \ref{sec:synthesis}).
%
%---to predict the distribution of QuickCheck
%generators\cite{DBLP:conf/haskell/MistaRH18}.
%
We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
automatically synthesize \quickcheck generators for \emph{rich ADT values},
where the distributions of random values can be adjusted at compile-time to what
developers might want based on our predictions.
%
Finally, we provide empirical evaluations showing that including static
information from the user codebase improves the code coverage of two external
applications when tested using random values generated following our ideas
(Section \ref{sec:casestudies}).

>>>>

% Scenario of random testing
Random testing is a promising approach to finding\todoale{find?} bugs
\cite{HughesNSA16,HughesPAN16,ArtsHNS15}.
%
\quickcheck \cite{ClaessenH00} is the dominant tool of this sort used by the
Haskell community.
%
It requires developers to specify (i) \emph{testing properties} describing
programs expected behavior and (ii) \emph{random data generators} based on the
\emph{types} of the expected inputs (e.g., an integer, an string,
etc.). %, e.g., an integer, an string, or an abstract data type (ADT).
%
\quickcheck then generates random test cases and reports violating testing
properties.

% Situation with Quick Check, ADTs, tools
\quickcheck comes equipped with random generators for built-in types, while it
requires to manually write generators for user-defined ADTs.
%
Recently, there have been a proliferation of tools to automatically derive
\quickcheck generators for ADTs
\cite{mitchell2007,RuncimanNL08,DuregardJW12,grieco2017,DBLP:conf/haskell/MistaRH18}.
%
The main difference about these tools lies on the guarantees provided to ensure
\emph{the termination of the generation process} and the \emph{distribution of
  random values}.
%
Despite their differences, these tools guarantee that generated values are
\emph{well-typed}.
%
In other words, generated values follow the structure described by ADT
definitions.

%% How to use random generated ADTs
Well-typed ADT values are specially useful when testing programs which expect
highly structured inputs like compilers \cite{Palka11,MidtgaardJKNN17}.
%
Generating ADT values also proves fruitful when looking for vulnerabilities in
combination with fuzzers \cite{GriecoCB16,grieco2017}.
%
%% Problem
%
Despite these success stories, ADT type-definitions do not often capture all the
invariants expected from the data that they are intended to model.
%
As a result, even if random values are well-typed, they might not be built with
enough structure to penetrate into deep software layers.

%% Our proposal
In this work, we identify two different sources of structural information that
can be statically exploited to improve the generation process of ADT values
(Section \ref{sec:sources}).
%
% In this work, we propose a novel improvement in the generation process of ADT
% values by exploiting some static information found in the codebase.
%
Then, we show how to capture this information into our (automatically) derived
random generators.
%
More specifically, we propose a generation process that is capable to consider
how programs branch on input ADTs values as well as how they get manipulated by
abstract interfaces (Section \ref{sec:hrepcont}).
%
Furthermore, we show how to predict the distribution of the \emph{expected}
numbers of ADT constructors, values fitting certain branching patterns, and
calls to interfaces.
%
For that, we extend some recent results on applying \emph{branching
  processes}\cite{gw1875}---a simple stochastic model conceived to study
population growth (Section \ref{sec:synthesis}).
%
%---to predict the distribution of QuickCheck
%generators\cite{DBLP:conf/haskell/MistaRH18}.
%
We implement our ideas in a tool called \dragenp\footnote{\dragenp is available
  at \url{http://github.com/OctopiChalmers/dragen-struct}}, that is capable to
automatically synthesize \quickcheck generators capable to produce \emph{rich ADT values},
where the distributions of random values can be adjusted at compile-time to what
developers might want. %based on our predictions.
%
Finally, we provide empirical evaluations showing that including static
information from the user codebase improves the code coverage of two external
applications when tested using random values generated following our ideas
(Section \ref{sec:casestudies}).
